---
title: "Block_preparation_fixedWindow"
author: "Goitom"
date: "2024-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This program will organize Haplotype block 
The blocks were constructed by PLINK according to fixed window size per chromosome average r2


Set the working directory where the files blocks_chr1.blocks to blocks_chr10.blocks located
```{r}
setwd()
```

The following steps will prepare the blocks for each chromosome 


Chromosome-1
```{r}
# Read the PLINK .block file into R without a header
block_data_chr1 <- read.delim("blocks_chromosome1.blocks", header = FALSE, comment.char = "", stringsAsFactors = FALSE)
dim(block_data_chr1) #2897 blocks
View(block_data_chr1)
# Split the characters in each cell and create a new data frame
block_data_str_chr1 <- data.frame(
  block = rep(1:length(block_data_chr1$V1), sapply(strsplit(block_data_chr1$V1, " "), length)),
  marker = unlist(strsplit(block_data_chr1$V1, " "))
)
head(block_data_str_chr1)
# Remove rows with asterisk (*) character in the Characters column
block_data_str_filt_chr1 <- subset(block_data_str_chr1, !grepl("\\*", marker))
head(block_data_str_filt_chr1)
str(block_data_str_filt_chr1)
block_data_str_filt_chr1$block <- as.character(block_data_str_filt_chr1$block)
head(block_data_str_filt_chr1)
str(block_data_str_filt_chr1)
```

Chromosome-2
```{r}
# Read the PLINK .block file into R without a header
block_data_chr2 <- read.delim("blocks_chromosome2.blocks", header = FALSE, comment.char = "", stringsAsFactors = FALSE)
dim(block_data_chr2) #2292
View(block_data_chr2)
# Split the characters in each cell and create a new data frame
block_data_str_chr2 <- data.frame(
  block = rep(1:length(block_data_chr2$V1), sapply(strsplit(block_data_chr2$V1, " "), length)),
  marker = unlist(strsplit(block_data_chr2$V1, " "))
)
head(block_data_str_chr2)
# Remove rows with asterisk (*) character in the Characters column
block_data_str_filt_chr2 <- subset(block_data_str_chr2, !grepl("\\*", marker))
head(block_data_str_filt_chr2)
str(block_data_str_filt_chr2)

#check the end of chromosome 1 block number to continue to block 1
tail(block_data_str_filt_chr1) # chr1 block ended at block 2893 2898
#then add that to the next chromosome block as continuation of blocks partitioning
block_data_str_filt_chr2$block <- block_data_str_filt_chr2$block + 
  as.numeric(block_data_str_filt_chr1$block[length(block_data_str_filt_chr1$block)])

block_data_str_filt_chr2$block <- as.character(block_data_str_filt_chr2$block)
head(block_data_str_filt_chr2)
str(block_data_str_filt_chr2)
```


Chromosome-3
```{r}
# Read the PLINK .block file into R without a header
block_data_chr3 <- read.delim("blocks_chromosome3.blocks", header = FALSE, comment.char = "", stringsAsFactors = FALSE)
dim(block_data_chr3) #2155
View(block_data_chr3)
# Split the characters in each cell and create a new data frame
block_data_str_chr3 <- data.frame(
  block = rep(1:length(block_data_chr3$V1), sapply(strsplit(block_data_chr3$V1, " "), length)),
  marker = unlist(strsplit(block_data_chr3$V1, " "))
)
head(block_data_str_chr3)
# Remove rows with asterisk (*) character in the Characters column
block_data_str_filt_chr3 <- subset(block_data_str_chr3, !grepl("\\*", marker))
head(block_data_str_filt_chr3)
str(block_data_str_filt_chr3)

#check the end of chromosome 2 block number to continue to block number of next chromosome
tail(block_data_str_filt_chr2) # chr2 block ended at block 5163 5166
#then add that to the next chromosome block as continuation of blocks partitioning
block_data_str_filt_chr3$block <- block_data_str_filt_chr3$block + 
  as.numeric(block_data_str_filt_chr2$block[length(block_data_str_filt_chr2$block)])

block_data_str_filt_chr3$block <- as.character(block_data_str_filt_chr3$block)
head(block_data_str_filt_chr3)
str(block_data_str_filt_chr3)
```

Chromosome-4
```{r}
# Read the PLINK .block file into R without a header
block_data_chr4 <- read.delim("blocks_chromosome4.blocks", header = FALSE, comment.char = "", stringsAsFactors = FALSE)
dim(block_data_chr4) #1736
View(block_data_chr4)
# Split the characters in each cell and create a new data frame
block_data_str_chr4 <- data.frame(
  block = rep(1:length(block_data_chr4$V1), sapply(strsplit(block_data_chr4$V1, " "), length)),
  marker = unlist(strsplit(block_data_chr4$V1, " "))
)
head(block_data_str_chr4)
# Remove rows with asterisk (*) character in the Characters column
block_data_str_filt_chr4 <- subset(block_data_str_chr4, !grepl("\\*", marker))
head(block_data_str_filt_chr4)
str(block_data_str_filt_chr4)

#check the end of chromosome 3 block number to continue to block number of next chromosome
tail(block_data_str_filt_chr3) # chr3 block ended at block 7313 7295
#then add that to the next chromosome block as continuation of blocks partitioning
block_data_str_filt_chr4$block <- block_data_str_filt_chr4$block + 
  as.numeric(block_data_str_filt_chr3$block[length(block_data_str_filt_chr3$block)])

block_data_str_filt_chr4$block <- as.character(block_data_str_filt_chr4$block)
head(block_data_str_filt_chr4)
str(block_data_str_filt_chr4)

```

Chromosome-5
```{r}
# Read the PLINK .block file into R without a header
block_data_chr5 <- read.delim("blocks_chromosome5.blocks", header = FALSE, comment.char = "", stringsAsFactors = FALSE)
dim(block_data_chr5) #2184
View(block_data_chr5)
# Split the characters in each cell and create a new data frame
block_data_str_chr5 <- data.frame(
  block = rep(1:length(block_data_chr5$V1), sapply(strsplit(block_data_chr5$V1, " "), length)),
  marker = unlist(strsplit(block_data_chr5$V1, " "))
)
head(block_data_str_chr5)
# Remove rows with asterisk (*) character in the Characters column
block_data_str_filt_chr5 <- subset(block_data_str_chr5, !grepl("\\*", marker))
head(block_data_str_filt_chr5)
str(block_data_str_filt_chr5)

#check the end of chromosome 4 block number to continue to block number of next chromosome
tail(block_data_str_filt_chr4) # chr4 block ended at block 9056 9034
#then add that to the next chromosome block as continuation of blocks partitioning
block_data_str_filt_chr5$block <- block_data_str_filt_chr5$block + 
  as.numeric(block_data_str_filt_chr4$block[length(block_data_str_filt_chr4$block)])

block_data_str_filt_chr5$block <- as.character(block_data_str_filt_chr5$block)
head(block_data_str_filt_chr5)
str(block_data_str_filt_chr5)

```

Chromosome-6
```{r}
# Read the PLINK .block file into R without a header
block_data_chr6 <- read.delim("blocks_chromosome6.blocks", header = FALSE, comment.char = "", stringsAsFactors = FALSE)
dim(block_data_chr6) #1492
View(block_data_chr6)
# Split the characters in each cell and create a new data frame
block_data_str_chr6 <- data.frame(
  block = rep(1:length(block_data_chr6$V1), sapply(strsplit(block_data_chr6$V1, " "), length)),
  marker = unlist(strsplit(block_data_chr6$V1, " "))
)
head(block_data_str_chr6)
# Remove rows with asterisk (*) character in the Characters column
block_data_str_filt_chr6 <- subset(block_data_str_chr6, !grepl("\\*", marker))
head(block_data_str_filt_chr6)
str(block_data_str_filt_chr6)

#check the end of chromosome 5 block number to continue to block number of next chromosome
tail(block_data_str_filt_chr5) # chr5 block ended at block 11246 11219
#then add that to the next chromosome block as continuation of blocks partitioning
block_data_str_filt_chr6$block <- block_data_str_filt_chr6$block + 
  as.numeric(block_data_str_filt_chr5$block[length(block_data_str_filt_chr5$block)])

block_data_str_filt_chr6$block <- as.character(block_data_str_filt_chr6$block)
head(block_data_str_filt_chr6)
str(block_data_str_filt_chr6)
```

Chromosome-7
```{r}
# Read the PLINK .block file into R without a header
block_data_chr7 <- read.delim("blocks_chromosome7.blocks", header = FALSE, comment.char = "", stringsAsFactors = FALSE)
dim(block_data_chr7) #1593
View(block_data_chr7)
# Split the characters in each cell and create a new data frame
block_data_str_chr7 <- data.frame(
  block = rep(1:length(block_data_chr7$V1), sapply(strsplit(block_data_chr7$V1, " "), length)),
  marker = unlist(strsplit(block_data_chr7$V1, " "))
)
head(block_data_str_chr7)
# Remove rows with asterisk (*) character in the Characters column
block_data_str_filt_chr7 <- subset(block_data_str_chr7, !grepl("\\*", marker))
head(block_data_str_filt_chr7)
str(block_data_str_filt_chr7)

#check the end of chromosome 6 block number to continue to block number of next chromosome
tail(block_data_str_filt_chr6) # chr6 block ended at block 12721 12677
#then add that to the next chromosome block as continuation of blocks partitioning
block_data_str_filt_chr7$block <- block_data_str_filt_chr7$block + 
  as.numeric(block_data_str_filt_chr6$block[length(block_data_str_filt_chr6$block)])

block_data_str_filt_chr7$block <- as.character(block_data_str_filt_chr7$block)
head(block_data_str_filt_chr7)
str(block_data_str_filt_chr7)
```


Chromosome-8
```{r}
# Read the PLINK .block file into R without a header
block_data_chr8 <- read.delim("blocks_chromosome8.blocks", header = FALSE, comment.char = "", stringsAsFactors = FALSE)
dim(block_data_chr8) #1567
View(block_data_chr8)
# Split the characters in each cell and create a new data frame
block_data_str_chr8 <- data.frame(
  block = rep(1:length(block_data_chr8$V1), sapply(strsplit(block_data_chr8$V1, " "), length)),
  marker = unlist(strsplit(block_data_chr8$V1, " "))
)
head(block_data_str_chr8)
# Remove rows with asterisk (*) character in the Characters column
block_data_str_filt_chr8 <- subset(block_data_str_chr8, !grepl("\\*", marker))
head(block_data_str_filt_chr8)
str(block_data_str_filt_chr8)

#check the end of chromosome 7 block number to continue to block number of next chromosome
tail(block_data_str_filt_chr7) # chr7 block ended at block 14300 14253
#then add that to the next chromosome block as continuation of blocks partitioning
block_data_str_filt_chr8$block <- block_data_str_filt_chr8$block + 
  as.numeric(block_data_str_filt_chr7$block[length(block_data_str_filt_chr7$block)])

block_data_str_filt_chr8$block <- as.character(block_data_str_filt_chr8$block)
head(block_data_str_filt_chr8)
str(block_data_str_filt_chr8)

```

Chromosome-9
```{r}
# Read the PLINK .block file into R without a header
block_data_chr9 <- read.delim("blocks_chromosome9.blocks", header = FALSE, comment.char = "", stringsAsFactors = FALSE)
dim(block_data_chr9) #1438 
View(block_data_chr9)
# Split the characters in each cell and create a new data frame
block_data_str_chr9 <- data.frame(
  block = rep(1:length(block_data_chr9$V1), sapply(strsplit(block_data_chr9$V1, " "), length)),
  marker = unlist(strsplit(block_data_chr9$V1, " "))
)
head(block_data_str_chr9)
# Remove rows with asterisk (*) character in the Characters column
block_data_str_filt_chr9 <- subset(block_data_str_chr9, !grepl("\\*", marker))
head(block_data_str_filt_chr9)
str(block_data_str_filt_chr9)

#check the end of chromosome 8 block number to continue to block number of next chromosome
tail(block_data_str_filt_chr8) # chr8 block ended at block 15856 15813
#then add that to the next chromosome block as continuation of blocks partitioning
block_data_str_filt_chr9$block <- block_data_str_filt_chr9$block + 
  as.numeric(block_data_str_filt_chr8$block[length(block_data_str_filt_chr8$block)])


block_data_str_filt_chr9$block <- as.character(block_data_str_filt_chr9$block)
head(block_data_str_filt_chr9)
str(block_data_str_filt_chr9)
```

Chromosome-10
```{r}
# Read the PLINK .block file into R without a header
block_data_chr10 <- read.delim("blocks_chromosome10.blocks", header = FALSE, comment.char = "", stringsAsFactors = FALSE)
dim(block_data_chr10) #1325 
View(block_data_chr10)
# Split the characters in each cell and create a new data frame
block_data_str_chr10 <- data.frame(
  block = rep(1:length(block_data_chr10$V1), sapply(strsplit(block_data_chr10$V1, " "), length)),
  marker = unlist(strsplit(block_data_chr10$V1, " "))
)
head(block_data_str_chr10)
# Remove rows with asterisk (*) character in the Characters column
block_data_str_filt_chr10 <- subset(block_data_str_chr10, !grepl("\\*", marker))
head(block_data_str_filt_chr10)
str(block_data_str_filt_chr10)

#check the end of chromosome 9 block number to continue to block number of next chromosome
tail(block_data_str_filt_chr9) # chr9 block ended at block 17281 17237
#then add that to the next chromosome block as continuation of blocks partitioning
block_data_str_filt_chr10$block <- block_data_str_filt_chr10$block + 
  as.numeric(block_data_str_filt_chr9$block[length(block_data_str_filt_chr9$block)])


block_data_str_filt_chr10$block <- as.character(block_data_str_filt_chr10$block)
head(block_data_str_filt_chr10)
str(block_data_str_filt_chr10)


```


Combine the blocks of each chromosome into a single data frame
```{r}
combined_blocks_chr1_10 <- rbind(block_data_str_filt_chr1 ,block_data_str_filt_chr2,block_data_str_filt_chr3,block_data_str_filt_chr4,
                                 block_data_str_filt_chr5,block_data_str_filt_chr6,block_data_str_filt_chr7,block_data_str_filt_chr8,
                                 block_data_str_filt_chr9,block_data_str_filt_chr10)
```

Write the formatted and combined block data to a new file
```{r}
write.table(combined_blocks_chr1_10, file = "all_chr_haplo_block.txt", sep = "\t", quote = FALSE, row.names = FALSE)

```


